{\em Cloudbreak}: Accurate and Scalable Genomic Structural Variation Detection in the Cloud with MapReduce

### 2014-04-30 manuscript/bmc_additional1.tex
drafts

\title{{\em Cloudbreak}: Accurate and Scalable Genomic Structural Variation Detection in the Cloud with MapReduce}

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{../figures/insert_size_mixtures.pdf}
\caption{Illustration of insert size mixtures at individual genomic locations. A) There is no variant present at the location indicated by the vertical line (left), so the mix of insert sizes (right) follows the expected distribution of the library centered at 200bp, with a small amount of noise coming from low-quality mappings. B) A homozygous deletion of 50bp at the location has shifted the distribution of observed insert sizes. C) A heterozygous deletion at the location causes a mixture of normal and long insert sizes to be detected. D) A heterozygous small insertion shifts a portion of the mixture to have lower insert sizes.}
\label{insert_size_mixes}
\end{figure}

@@

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{r|r|rr|rr|}
\multicolumn{2}{c}{}  & \multicolumn{4}{c}{Actual Genotypes} \\
\multicolumn{2}{c}{}  & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\cline{3-6}
\multicolumn{2}{c|}{} &  Homozygous & Heterozygous & Homozygous & Heterozygous \\ 
\cline{2-6}
\multirow{2}{*}{\shortstack{Predicted \\ Genotypes}} & Homozygous & 35 & 2 &  96 & 21 \\
 & Heterozygous & 0 & 39 &  2 & 448 \\
\cline{2-6}
\end{tabular}
\end{center}
\caption{Table A4. Confusion matrices for the predicted genotype of deletions found by Cloudbreak on both the simulated and NA18507 data sets.}
\label{deletionGenotypeaccuracy}
\end{table}

>>>>

\title{{\em Cloudbreak}: Accurate and Scalable Genomic Structural Variation Detection in the Cloud with MapReduce \\ (Appendix)}

@@

\tableofcontents

\section{Additional Figures}

\bigskip

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{../figures/insert_size_mixtures.pdf}
\caption{Illustration of insert size mixtures at individual genomic locations. A) There is no variant present at the location indicated by the vertical line (left), so the mix of insert sizes (right) follows the expected distribution of the library centered at 200bp, with a small amount of noise coming from low-quality mappings. B) A homozygous deletion of 50bp at the location has shifted the distribution of observed insert sizes. C) A heterozygous deletion at the location causes a mixture of normal and long insert sizes to be detected. D) A heterozygous small insertion shifts a portion of the mixture to have lower insert sizes.}
\label{insert_size_mixes}
\end{figure}

@@

\section{Additional Tables}

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{r|r|rr|rr|}
\multicolumn{2}{c}{}  & \multicolumn{4}{c}{Actual Genotypes} \\
\multicolumn{2}{c}{}  & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\cline{3-6}
\multicolumn{2}{c|}{} &  Homozygous & Heterozygous & Homozygous & Heterozygous \\ 
\cline{2-6}
\multirow{2}{*}{\shortstack{Predicted \\ Genotypes}} & Homozygous & 35 & 2 &  96 & 21 \\
 & Heterozygous & 0 & 39 &  2 & 448 \\
\cline{2-6}
\end{tabular}
\end{center}
\caption{Confusion matrices for the predicted genotype of deletions found by Cloudbreak on both the simulated and NA18507 data sets.}
\label{deletionGenotypeaccuracy}
\end{table}

### 2014-04-30 manuscript/whelan_et_al_cloudbreak_ed.tex
drafts

Hadoop/MapReduce requires a specific programming model, however, which can make it difficult to design general-purpose algorithms for arbitrary sequencing data processing problems such as SV detection. MapReduce divides computation across a cluster into three phases. In the first phase, \emph{mappers} developed by the application programmer examine small blocks of data and emit a set of $\langle key, value \rangle$ pairs for each block examined. The framework then sorts the output of the mappers by key, and aggregates all values that are associated with each key. Finally, the framework executes \emph{reducers}, also created by the application developer, which process all of the values for a particular key and produce one or more outputs that summarize or aggregate those values.

@@

  \subsection*{Table 2 - Variants Detected in the Simulated Data Set}
The number of simulated deletions and insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those variants that were discovered exclusively by each tool (in parentheses). The total number of variants in each size class in the true set of deletions and insertions is shown in the first row of each section. \par \mbox{}
\par
\mbox{
\begin{tabular}{r|rrr|rrrrr}
  \cline{2-9}
   &                     & Prec. & Recall & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
\hline
\multirow{7}{*}{\begin{sideways}Deletions\end{sideways}} & Total Number &          &           & 224 &  84 & 82 &  31 & 26\\ 
  \hline
\cline{2-9}
&  Cloudbreak    &  0.638 & \textbf{0.678} & \textbf{153} (9)  & 61 (0) &  62 (0) & 12 (0) & 15 (0) \\ 
&  BreakDancer   &  0.356 & 0.49 & 89 (0)  & 54 (0) &  53 (0) & 8 (0) & 15 (0) \\ 
&  GASVPro        & 0.146 & 0.432 & 83 (2)  & 32 (0) &  55 (0) & 8 (0) & 15 (0) \\ 
&  DELLY-RP           & 0.457 & 0.613 & 114 (3)  & \textbf{68} (0) &  \textbf{66} (0) & 9 (1) & 17 (0) \\ 
&  DELLY-SR           & \textbf{0.679} & 0.166 & 0 (0)  & 3 (0) &  49 (0) & 6 (0) & 16 (0) \\ 
&  Pindel           & 0.462 & 0.421 & 96 (\textbf{11})  & 24 (0) &  48 (0) & 5 (0) & 15 (0)\\ 
&  MoDIL           & 0.132  & 0.66 & 123 (6)  & 66 (\textbf{3}) &  \textbf{66} (\textbf{11}) & \textbf{17} (\textbf{7}) & \textbf{23} (\textbf{8})\\ 
   \hline
\multirow{5}{*}{\begin{sideways}Insertions\end{sideways}} & Total Number &          &           & 199 &  83 & 79 &  21 & 21\\ 
\cline{2-9}
&  Cloudbreak   &0.451 & \textbf{0.305}  & \textbf{79} (\textbf{32})  & \textbf{32} (\textbf{18}) &  \textbf{11} (8) & 1 (0) & 0 (0) \\ 
&  BreakDancer & 0.262 & 0.0968  & 23 (5)  & 14 (5) &  2 (1) & 0 (0) & 0 (0) \\ 
&  Pindel          & \textbf{0.572} & 0.196 & 52 (25)  & 5 (1) &  10 (\textbf{9}) & \textbf{3} (\textbf{2}) & \textbf{9} (\textbf{9})\\ 
&  MoDIL          & 0.186 & 0.0521 & 14 (1)  & 4 (0) &  1 (0) & 2 (\textbf{2}) & 0 (0)\\ 
\hline
\end{tabular}
}

>>>>

Hadoop/MapReduce requires a specific programming model, however, which can make it difficult to design general-purpose algorithms for arbitrary sequence processing problems such as SV detection. MapReduce divides computation across a cluster into three phases. In the first phase, \emph{mappers} developed by the application programmer examine small blocks of data and emit a set of $\langle key, value \rangle$ pairs for each block examined. The framework then sorts the output of the mappers by key, and aggregates all values that are associated with each key. Finally, the framework executes \emph{reducers}, also created by the application developer, which process all of the values for a particular key and produce one or more outputs that summarize or aggregate those values.

@@

  \subsection*{Table 2 - Variants detected in the simulated data set}
The number of simulated deletions and insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those variants that were discovered exclusively by each tool (in parentheses). The total number of variants in each size class in the true set of deletions and insertions is shown in the first row of each section. \par \mbox{}
\par
\mbox{
\begin{tabular}{r|rrr|rrrrr}
  \cline{2-9}
   &                     & Prec. & Recall & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
\hline
\multirow{7}{*}{\begin{sideways}Deletions\end{sideways}} & Total Number &          &           & 224 &  84 & 82 &  31 & 26\\ 
  \hline
\cline{2-9}
&  Cloudbreak    &  0.638 & \textbf{0.678} & \textbf{153} (9)  & 61 (0) &  62 (0) & 12 (0) & 15 (0) \\ 
&  BreakDancer   &  0.356 & 0.49 & 89 (0)  & 54 (0) &  53 (0) & 8 (0) & 15 (0) \\ 
&  GASVPro        & 0.146 & 0.432 & 83 (2)  & 32 (0) &  55 (0) & 8 (0) & 15 (0) \\ 
&  DELLY-RP           & 0.457 & 0.613 & 114 (3)  & \textbf{68} (0) &  \textbf{66} (0) & 9 (1) & 17 (0) \\ 
&  DELLY-SR           & \textbf{0.679} & 0.166 & 0 (0)  & 3 (0) &  49 (0) & 6 (0) & 16 (0) \\ 
&  Pindel           & 0.462 & 0.421 & 96 (\textbf{11})  & 24 (0) &  48 (0) & 5 (0) & 15 (0)\\ 
&  MoDIL           & 0.132  & 0.66 & 123 (6)  & 66 (\textbf{3}) &  \textbf{66} (\textbf{11}) & \textbf{17} (\textbf{7}) & \textbf{23} (\textbf{8})\\ 
   \hline
\multirow{5}{*}{\begin{sideways}Insertions\end{sideways}} & Total Number &          &           & 199 &  83 & 79 &  21 & 21\\ 
\cline{2-9}
&  Cloudbreak   &0.451 & \textbf{0.305}  & \textbf{79} (\textbf{32})  & \textbf{32} (\textbf{18}) &  \textbf{11} (8) & 1 (0) & 0 (0) \\ 
&  BreakDancer & 0.262 & 0.0968  & 23 (5)  & 14 (5) &  2 (1) & 0 (0) & 0 (0) \\ 
&  Pindel          & \textbf{0.572} & 0.196 & 52 (25)  & 5 (1) &  10 (\textbf{9}) & \textbf{3} (\textbf{2}) & \textbf{9} (\textbf{9})\\ 
&  MoDIL          & 0.186 & 0.0521 & 14 (1)  & 4 (0) &  1 (0) & 2 (\textbf{2}) & 0 (0)\\ 
\hline
\end{tabular}
}
